---
title: "HW4"
author: "Vishnu Rahul"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(dplyr)
library(readxl)
library(cluster)
library(ggplot2)
library(tidyr)
```

```{r fileimport, include=FALSE}
prospects <- read_excel("~/Downloads/prospect.xls")
```

#exclude the fields LOCATION and ID
```{r exclude, include=TRUE}
prospects$ID <- NULL
prospects$LOC <- NULL 
```

#Check for NA
```{r NA, include=TRUE}
any_is_na = apply(prospects, 2, function(x) any(is.na(x)))
any_is_na
sum(any_is_na)
```

#Explore the data
```{r Explore, include=TRUE}
# plot histogram to see spread of interval data
hist(prospects$AGE)
hist(prospects$INCOME)
hist(prospects$`FICO>=700`)

#get count of boolean and ordinal values
prospects %>% 
  group_by(SEX) %>%
  summarise(no_rows = length(SEX))
prospects %>% 
  group_by(MARRIED) %>%
  summarise(no_rows = length(MARRIED))
prospects %>% 
  group_by(OWNHOME) %>%
  summarise(no_rows = length(OWNHOME))
prospects %>% 
  group_by(CLIMATE) %>%
  summarise(no_rows = length(CLIMATE))
prospects %>% 
  group_by(`FICO>=700`) %>%
  summarise(no_rows = length(`FICO>=700`))

# convert NA values in SEX to "M" because it is the majority class
prospects$SEX[is.na(prospects$SEX)] <- "M"

# convert "M" to 1 and "F" to 0
prospects$SEX[prospects$SEX == "M"] <- 1
prospects$SEX[prospects$SEX == "F"] <- 0

# convert NA values in MARRIED to 1 because it is the majority class
prospects$MARRIED[is.na(prospects$MARRIED)] <- 1

# convert NA values in OWNHOME to 0 because it is the majority class
prospects$OWNHOME[is.na(prospects$OWNHOME)] <- 0

# convert NA values in FICO>=700 to 1 because it is the majority class
prospects$`FICO>=700`[is.na(prospects$`FICO>=700`)] <- 1

# convert NA values in AGE and INCOME to the mean of the respective attribute
prospects$AGE[is.na(prospects$AGE)] = mean(prospects$AGE, na.rm = T)
prospects$INCOME[is.na(prospects$INCOME)] = mean(prospects$INCOME, na.rm = T)

# convert all the boolean/ordinal attributes into a factor
prospects$CLIMATE <- as.factor(prospects$CLIMATE)
prospects$OWNHOME <- as.factor(prospects$OWNHOME)
prospects$MARRIED <- as.factor(prospects$MARRIED)
prospects$SEX <- as.factor(prospects$SEX)
```
#Check for NA
```{r NA, include=TRUE}
any_is_na = apply(prospects, 2, function(x) any(is.na(x)))
any_is_na
sum(any_is_na)
```
#Construct k means model with k=4
```{r kmeans, include=TRUE}
set.seed(7)
km <- kmeans(prospects, 4, nstart = 100)

#Number of points in each cluster
km$size

#cluster means
km$centers

#variances within each cluster
km$withinss

#variances between each cluster (between is much higher than within => clusters are better)
km$betweenss

#quality of the cluster
km$betweenss/km$totss 
# Within cluster sum of squares by cluster: 63%

```
#Cluster Characteristics
```{r kmeans, include=TRUE}
cluster_one   <- prospects[km$cluster == 1,]
cluster_two   <- prospects[km$cluster == 2,]
cluster_three <- prospects[km$cluster == 3,]
cluster_four  <- prospects[km$cluster == 4,]


cluster_one %>% 
  group_by(SEX) %>%
  summarise(no_rows = length(SEX))
cluster_one %>% 
  group_by(MARRIED) %>%
  summarise(no_rows = length(MARRIED))
cluster_one %>% 
  group_by(OWNHOME) %>%
  summarise(no_rows = length(OWNHOME))
cluster_one %>% 
  group_by(CLIMATE) %>%
  summarise(no_rows = length(CLIMATE))
cluster_one %>% 
  group_by(`FICO>=700`) %>%
  summarise(no_rows = length(`FICO>=700`))


cluster_two %>% 
  group_by(SEX) %>%
  summarise(no_rows = length(SEX))
cluster_two %>% 
  group_by(MARRIED) %>%
  summarise(no_rows = length(MARRIED))
cluster_two %>% 
  group_by(OWNHOME) %>%
  summarise(no_rows = length(OWNHOME))
cluster_two %>% 
  group_by(CLIMATE) %>%
  summarise(no_rows = length(CLIMATE))
cluster_two %>% 
  group_by(`FICO>=700`) %>%
  summarise(no_rows = length(`FICO>=700`))

cluster_three %>% 
  group_by(SEX) %>%
  summarise(no_rows = length(SEX))
cluster_three %>% 
  group_by(MARRIED) %>%
  summarise(no_rows = length(MARRIED))
cluster_three %>% 
  group_by(OWNHOME) %>%
  summarise(no_rows = length(OWNHOME))
cluster_three %>% 
  group_by(CLIMATE) %>%
  summarise(no_rows = length(CLIMATE))
cluster_three %>% 
  group_by(`FICO>=700`) %>%
  summarise(no_rows = length(`FICO>=700`))

cluster_four %>% 
  group_by(SEX) %>%
  summarise(no_rows = length(SEX))
cluster_four %>% 
  group_by(MARRIED) %>%
  summarise(no_rows = length(MARRIED))
cluster_four %>% 
  group_by(OWNHOME) %>%
  summarise(no_rows = length(OWNHOME))
cluster_four %>% 
  group_by(CLIMATE) %>%
  summarise(no_rows = length(CLIMATE))
cluster_four %>% 
  group_by(`FICO>=700`) %>%
  summarise(no_rows = length(`FICO>=700`))


cluster_one %>%
  gather(attributes, value, 1:2) %>%
  ggplot(aes(x = value)) +
  geom_histogram(fill = 'lightblue2', color = 'black') +
  facet_wrap(~attributes, scales = 'free_x') +
  labs(x="Values", y="Frequency") +
  theme_bw()

cluster_two %>%
  gather(attributes, value, 1:2) %>%
  ggplot(aes(x = value)) +
  geom_histogram(fill = 'lightblue2', color = 'black') +
  facet_wrap(~attributes, scales = 'free_x') +
  labs(x="Values", y="Frequency") +
  theme_bw()

cluster_three %>%
  gather(attributes, value, 1:2) %>%
  ggplot(aes(x = value)) +
  geom_histogram(fill = 'lightblue2', color = 'black') +
  facet_wrap(~attributes, scales = 'free_x') +
  labs(x="Values", y="Frequency") +
  theme_bw()

cluster_four %>%
  gather(attributes, value, 1:2) %>%
  ggplot(aes(x = value)) +
  geom_histogram(fill = 'lightblue2', color = 'black') +
  facet_wrap(~attributes, scales = 'free_x') +
  labs(x="Values", y="Frequency") +
  theme_bw()


# Cluster 1
# Mostly males around the age of 30-40 and income around 50k-60k; who are un-married, who dont own a home, live in climate 20 and fico score is less than 700


# Cluster 2
# Roughly equal number of male and female around the age of 40-53 and income around 38k - 50k; who are married, who dont own a home, live in climate 20 and fico score is almost equally distributed between FICO_Score < 700 and FICO_Score>=700


# Cluster 3
# Mostly males around the age of 45-55 and income around 58k-66k; who are married, who dont own a home, live in climate 20 and fico score is greater than 700


# Cluster 4
# Mostly Female with age around 30 to 50 and income around 15k - 35k who are married, who dont own a home, live in climate 20 and fico score is less than 700
```  
  
#plot the clusters NA KM
```{r km plot, include=TRUE}
#Cluster visualization for Age V/s Income
plot(prospects[,1:2],col=(km$cluster))

#Overall customer visualization
plot(prospects,col=(km$cluster))
```
##Elbow Method for finding the optimal number of clusters
```{r elbow plot, include=TRUE}
#Elbow Method for finding the optimal number of clusters
set.seed(123)
# Compute and plot wss for k = 2 to k = 15.
k.max <- 15
data <- prospects
wss <- sapply(1:k.max, 
              function(k){kmeans(data, k, nstart=100)$tot.withinss})
#plot wss
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")

#we choose k= 6
```
#rerun kmeans model with optimal k = 6
```{r rerun, include=TRUE}
km1 <- kmeans(prospects, 6, nstart = 100)
plot(prospects[,1:2],col=(km1$cluster))

km1
#we can see that Within cluster sum of squares by cluster improved to 71.5 %
```

# Check Silhouette measure for k = 6 (if it closes to 1, means good)
```{r Sil, include=TRUE}
library("cluster")

avg_sil <- function(k){
  Data <- prospects
  kmModel <- kmeans(Data,centers = k, nstart = 100)
  ss <- silhouette(kmModel$cluster, dist(data))
  mean(ss[,3])
}

avg_sil(6)

# the average Silhouette measure = 0.2522056 which is good 
```

#rerun kmeans model with optimal k = 8
```{r rerun2, include=TRUE}
km2 <- kmeans(prospects, 8, nstart = 100)
km2
plot(prospects[,1:2],col=(km2$cluster))
#we can see that Within cluster sum of squares by cluster improved to 
# (between_SS / total_SS =  76.5 %)
```

# Check Silhouette measure for k = 8 
```{r Sil2, include=TRUE}
avg_sil(8)

# the average Silhouette measure = 0.2481758 which is still good 

# Based on theWithin cluster sum of squares by cluster and Silhouette measure, we choose k=8
```